---
title: "Day 2D Activity<br />(Resampling Example)"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    df_print: paged
editor_options: 
  chunk_output_type: inline
---


```{r, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  df_print = "paged"
)
```

# Load packages

```{r, message=FALSE}
# Set all of this to get the EXACT SAME results on all platforms
set.seed(2022, "Mersenne-Twister", "Inversion", "Rejection")

library(tidyverse)
library(tidymodels)
library(vip)
library(palmerpenguins)
penguins <- penguins %>% na.omit()
```

# Hands-on Activity

Our goal is to build a model predicting `body_mass_g`. 

1. Create an initial split, stratified by the label.

2. Create a recipe:
      a. Create dummy variables for categorical features 
      b. Normalize all numeric features
      c. Remove features with near-zero variance
      d. Remove features with large absolute correlations with other features
      e. Remove features that are a linear combination of other features

<p style="padding-top:0.01px;">3. Specify a linear regression model 

4. Build a workflow

5. Fit a model on the training data with 10-fold cross-validation, repeated three times, stratified by the label

6. Examine performance during cross-validation 

7. Fit a final model on the training data and plot variable importance

8. Evaluate performance in the final, hold-out test set. 

---
 
## Answer key

<details><summary>Click here to view the answer key to the hands-on activity</summary>

```{r}
# 1
bmg_split <- initial_split(penguins, prop = 0.8, strata = body_mass_g)
bmg_train <- training(bmg_split)
bmg_test <- testing(bmg_split)
```

```{r}
# 2
bmg_recipe <- 
  recipe(bmg_train, formula = body_mass_g ~ .) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_nzv(all_predictors()) %>%
  step_corr(all_predictors()) %>%
  step_lincomb(all_predictors()) %>% 
  step_normalize(all_numeric_predictors())
```

```{r}
# 3 
lin_reg <- linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")

# 4
bmg_wflow <- 
  workflow() %>% 
  add_model(lin_reg) %>% 
  add_recipe(bmg_recipe)
```

```{r}
#5
bmg_folds <- vfold_cv(
  data = bmg_train,
  v = 10, 
  repeats = 3,
  strata = body_mass_g
)

bmg_fitr <- 
  bmg_wflow %>%
  fit_resamples(
    resamples = bmg_folds
  )
```

```{r}
# 6
collect_metrics(bmg_fitr)

# 7
bmg_final <- 
  bmg_wflow %>% 
  last_fit(bmg_split)

bmg_final %>% extract_fit_parsnip() %>% vip()
```

```{r}
# 8. 
collect_metrics(bmg_final)

bmg_pred <- 
  bmg_final %>% 
  collect_predictions() 

bmg_pred %>%
  ggplot(aes(x = body_mass_g, y = .pred)) + 
  geom_point(alpha = .15) + 
  geom_abline(color = 'darkred') + 
  coord_obs_pred() + 
  labs(x = "Observed", 
       y = "Predicted")
```

</details>
