---
title: '<span style="font-size:48pt;">Logistics & Data Exploration</span>'
subtitle: '&#x1F4CB; &#x1F6E0; &#x1F4CA;' 
author: 'Applied Machine Learning in R </br> Pittsburgh Summer Methodology Series'
date: 'Lecture 1-B &emsp; &emsp; July 19, 2021'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current% / %total%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#23395b"
)
```

```{r packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(countdown)
library(patchwork)
```

class: inverse, center, middle
# Packages
---
## caret

The `caret` (**C**lassification **A**nd **RE**gression **T**raining) package includes functions for:
- data visualization
- data pre-processing
- feature selection
- data splitting
- model training & testing
- variable importance estimation

--

<p style="padding-top:20px;"> There are many packages for building machine learning models, which often have different syntax.

`caret` <span style="color:darkred;">streamlines and standardizes</span> the process for creating predictive models.

It includes 50+ machine learning models.

It has historically been the <span style="color:darkred;">most popular package</span> for machine learning in R, with many resources, solutions, and answers to questions available online.

---
## tidymodels

The `tidymodels` meta-package is a collection of many packages for machine learning, including: 
- `rsample` for data splitting and resampling
- `recipes` for pre-processing
- `parsnip` for trying out many models
- `workflows` to streamline the pre-processing, modeling, and post-processing
- `tune` to optimize model hyperparameters
- `yardstick` for model performance metrics
- `broom` for converting information to user-friendly formats
- `dials` for creating and managing tuning parameters

--

<p style="padding-top:20px;"> Both `tidymodels` and `caret` are written by the same author (Max Kuhn). 

While `caret` has been around for a long time, `tidymodels` is much newer and built with <span style="color:darkred;">`tidyverse`</span> principles.

---
## Packages

We will demonstrate use of both modeling frameworks (`caret` and `tidyverse`) throughout the course.

You are free to use whichever framework you prefer during the small group and individual activities. 

--

<p style="padding-top:20px;">Please make sure you have both packages installed: 

`install.packages("caret")`

`install.packages("tidymodels")`

---
class: inverse, center, middle
# Exploratory Data Analysis

---
## Why exploratory data analysis?

Provide insight into your dataset

Determine whether your data are appropriate for the task

Detect data anamolies, outliers, and missing data

Explore nonlinearities in data

Help inform feature selection



