---
title: "Course Datasets"
output: 
  html_document:
    toc: FALSE
    df_print: paged
    self-contained: true
---

The following datasets will be used in the course:

Name | Description | Type | Link
:--- | :---------- | :--- | :---
titanic | Data on passengers of the Titanic | Real | https://bit.ly/amlr-titanic
airsat | Data on Airline customer satisfaction | Real | https://bit.ly/amlr-airsat
iris | Data on flower measurements | Real | https://bit.ly/amlr-iris
globalWarm | Data on emotions, ideology, and attitudes towards action on climate change | Real | https://bit.ly/amlr-globalWarm
heart | Data on heart disease predictions | Real | https://bit.ly/amlr-heart
ahd | Data on Alzheimer's disease | Real | https://bit.ly/amlr-ahd
hcp_memory | Neuroimaging data and memory tests | Modified from real | https://bit.ly/amlr-memory

---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```

## titanic

Real data on 963 passengers on the Titanic

- Short URL: https://bit.ly/amlr-titanic
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/titanic.csv

Variables

- `survived`: Did the passenger survive the disaster? {no, yes}
- `pclass`: Which class did the passenger travel in? {1st, 2nd, 3rd}
- `sex`: What was the passenger's sex? {female, male}
- `age`: What was the passenger's age in years? (0.1667, 80.0000)
- `sibsp`: How many of the passenger's siblings and spouses were aboard? (0, 8)
- `parch`: How many of the passenger's parents or children were aboard? (0, 6)
- `fare`: How much did the passenger pay for transportation in dollars? (4.0125, 512.3292)

```{r, echo=FALSE}
read_csv("https://bit.ly/amlr-titanic", col_types = "fffdiid")
```

---

## airsat

Real data on 10,000 customers of an airline

- Short URL: https://bit.ly/amlr-airsat
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/airsatisfaction.csv

Variables

- `satisfaction`: Was the customer satisfied? {dissatisfied, satisfied}
- `sex`: What was the customer's sex? {Female, Male}
- `customer_type`: What type of customer are they? {disloyal, loyal}
- `age`: How old is the customer in years? (7, 85)
- `travel_type`: What type of trip was this? {Business, Personal}
- `class`: What class did the customer travel in? {Business, Eco, Eco Plus}
- `flight_distance`: How far was the flight in kilometers? (50, 6907)
- `seat_comfort`:`online_boarding`: Level of satisfaction with aspects of trip {1, 2, 3, 4, 5} (0=NA)
- `departure_delay`: How many minutes was departure delayed? (0, 569)
- `arrival_delay`: How many minutes was arrival delayed? (0, 543)

```{r, echo=FALSE}
read_csv("https://bit.ly/amlr-airsat", col_types = "fffdffdiiiiiiiiiiiiiidd")
```

---

## iris

Real data on 150 iris flowers. 

- Available from the base R {datasets} package. 

Variables

- `Sepal.Length`: length of flower sepal in cm (4.3, 7.9)
- `Sepal.Width`: width of flower sepal in cm (2.0, 4.4)
- `Petal.Length`: length of flower petal in cm (1.0, 6.9)
- `Petal.Width`: width of flower petal in cm (0.1, 2.5)
- `Species`: type of iris flower {setosa, versicolor, virginica}

```{r, echo = FALSE}
data(iris)
iris %>% as_tibble()
```

---

## globalWarm

Real data on emotions, ideology, and party affiliation as predictors of attitudes towards government action on climate change. 

- Available from the {psychTools} package. 

Variables 

- `govact`: attitudes towards government action on climate change (1, 7)
- `posemot`: positive emotions about climate change (1, 6)
- `negemot`: negative emotions about climate change (1, 6)
- `ideology`: political ideology/conservatism (1, 7)
- `age`: participant age (17, 87)
- `sex`: participant sex {female, male}
- `partyid`: political party {democrat, independent, republican}

---

## heart 

Real data on risk for heart disease. 

- Short URL: https://bit.ly/amlr-heart
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/heart.csv

```{r, echo = FALSE}
read_csv("https://bit.ly/amlr-heart", col_types = "dffddffdfdiiif")
```

---

## ahd 

Real data on Alzheimer's Disease. 

- Short URL: https://bit.ly/amlr-ahd
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/ahd.csv

```{r, echo = FALSE}
read_csv("https://bit.ly/amlr-ahd", col_types = list(male = "f", Genotype = "f", diagnosis = "f"))
```

---

## hcp_memory

Real neuroimaging data from the Human Connectome Project used to predict scores on a memory test. Note: artificially modified to increase predictive power and make activities more engaging.  

- Short URL: https://bit.ly/amlr-memory
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/hcp_memory.csv

Variables

```{r, echo = FALSE}
read_csv("https://bit.ly/amlr-memory")
```


