---
title: "Course Datasets"
output: 
  html_document:
    toc: FALSE
    df_print: paged
    self-contained: true
---

The following datasets will be used in the course:

Name | Description | Type | Link
:--- | :---------- | :--- | :---
titanic | Data on passengers of the Titanic | Real | https://bit.ly/amlr-titanic
airsat | Data on Airline customer satisfaction | Real | https://bit.ly/amlr-airsat

---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```

## titanic

Real data on 963 passengers on the Titanic

- Short URL: https://bit.ly/amlr-titanic
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/titanic.csv

Variables

- `survived`: Did the passenger survive the disaster? {no, yes}
- `pclass`: Which class did the passenger travel in? {1st, 2nd, 3rd}
- `sex`: What was the passenger's sex? {female, male}
- `age`: What was the passenger's age in years? (0.1667, 80.0000)
- `sibsp`: How many of the passenger's siblings and spouses were aboard? (0, 8)
- `parch`: How many of the passenger's parents or children were aboard? (0, 6)
- `fare`: How much did the passenger pay for transportation in dollars? (4.0125, 512.3292)

```{r, echo=FALSE}
read_csv("https://bit.ly/amlr-titanic", col_types = "fffdiid")
```

---

## airsat

Real data on 10,000 customers of an airline

- Short URL: https://bit.ly/amlr-airsat
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/airsatisfaction.csv

Variables

- `satisfaction`: Was the customer satisfied? {dissatisfied, satisfied}
- `sex`: What was the customer's sex? {Female, Male}
- `customer_type`: What type of customer are they? {disloyal, loyal}
- `age`: How old is the customer in years? (7, 85)
- `travel_type`: What type of trip was this? {Business, Personal}
- `class`: What class did the customer travel in? {Business, Eco, Eco Plus}
- `flight_distance`: How far was the flight in kilometers? (50, 6907)
- `seat_comfort`:`online_boarding`: Level of satisfaction with aspects of trip {1, 2, 3, 4, 5} (0=NA)
- `departure_delay`: How many minutes was departure delayed? (0, 569)
- `arrival_delay`: How many minutes was arrival delayed? (0, 543)

```{r, echo=FALSE}
read_csv("https://bit.ly/amlr-airsat", col_types = "fffdffdiiiiiiiiiiiiiidd")
```

