---
title: "Course Datasets"
output: 
  html_document:
    toc: FALSE
    df_print: paged
    self-contained: true
---

The following datasets will be used in the course:

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```

## airsat

Real data on the satisfaction and experience of 10,000 customers of an airline <a href="https://www.kaggle.com/datasets/teejmahal20/airline-passenger-satisfaction" class="badge badge-info">Info</a>

  -   Classification task: predict `satisfaction` (4A)

```{r, eval=FALSE}
# install.packages("readr")
airsat <- readr::read_csv("https://tinyurl.com/mlr-airsat")
```

```{r, echo=FALSE, message=FALSE}
# https://github.com/SMaRTWorkshops/mlr/raw/main/data/airsat.csv
read_csv("./data/airsat.csv", show_col_types = FALSE)
```


---

## ema

Real ecological momentary assessment (EMA) data on 80 undergraduate students in the Netherlands during the start of COVID (from Fried, Papanikolaou, & Epskamp, 2021) <a href="https://eiko-fried.com/wp-content/uploads/21677026211017839.pdf" class="badge badge-info">Info</a>

  - Data splitting for multilevel and time series data (1C)
  
```{r, eval=FALSE}
# install.packages("readr")
ema <- readr::read_csv("https://tinyurl.com/mlr-ema", col_types = cols(id = "f"))
```

```{r, echo=FALSE, message=FALSE}
# https://raw.githubusercontent.com/SMaRTWorkshops/mlr/main/data/ema.csv
readr::read_csv("./data/ema.csv", col_types = cols(id = "f"), show_col_types = FALSE)
```

---

## penguins

Real data on 344 penguins observed at Palmer Station, Antarctica <a href="https://allisonhorst.github.io/palmerpenguins/" class="badge badge-info">Info</a>

  - Regression task: predict `body_mass_g` (1D)

```{r}
modeldata::penguins
```

---

## titanic

Real data on 1,309 passengers on the Titanic <a href="https://www.openml.org/search?type=data&sort=runs&id=40945&status=active" class="badge badge-info">Info</a>

  - Classification task: predict `survived` (1D, 2A, 2D)
  
  - Regression task: predict `fare` (2A, 2B, 2C)

```{r, eval=FALSE}
# install.packages("readr")
titanic <- readr::read_csv("https://tinyurl.com/mlr-titanic")
```

```{r, echo=FALSE, message=FALSE}
# https://raw.githubusercontent.com/SMaRTWorkshops/mlr/main/data/titanic.csv
read_csv("./data/titanic.csv", show_col_types = FALSE)
```

---

## water

Real data on the potability and chemical properties of 2,011 water bodies <a href="https://www.kaggle.com/adityakadiwal/water-potability" class="badge badge-info">Info</a>

  - Classification task: predict `Potability` (3B, 3D)
  
```{r, eval=FALSE}
# install.packages("readr")
water <- readr::read_csv("https://tinyurl.com/mlr-water")
```

```{r, echo=FALSE, message=FALSE}
# https://raw.githubusercontent.com/SMaRTWorkshops/mlr/main/data/water.csv
read_csv("./data/water.csv", show_col_types = FALSE)
```
