---
title: "Course Datasets"
output: 
  html_document:
    toc: FALSE
    df_print: paged
    self-contained: true
---

The following datasets will be used in the course:

---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```

## titanic

Real data on 963 passengers on the Titanic

- Short URL: https://bit.ly/amlr-titanic
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/titanic.csv

```{r, echo=FALSE, message=FALSE}
read_csv("https://bit.ly/amlr-titanic", col_types = "fffdiid")
```

---

## airsat

Real data on 10,000 customers of an airline

- Short URL: https://bit.ly/amlr-airsat
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/airsatisfaction.csv

```{r, echo=FALSE, message=FALSE}
read_csv("https://bit.ly/amlr-airsat", col_types = "fffdffdiiiiiiiiiiiiiidd")
```

---

## iris

Real data on 150 iris flowers. 

- Available from the base R {datasets} package. 

```{r, echo = FALSE, message=FALSE}
data(iris)
iris %>% as_tibble()
```

---

## globalWarm

Real data on emotions, ideology, and party affiliation as predictors of attitudes towards government action on climate change. 

- Available from the {psychTools} package. 

```{r, echo=FALSE, message=FALSE}
data("globalWarm", package = "psychTools")
globalWarm
```


---

## heart 

Real data on risk for heart disease. 

- Short URL: https://bit.ly/amlr-heart
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/heart.csv

```{r, echo = FALSE, message=FALSE}
read_csv("https://bit.ly/amlr-heart", col_types = "dffddffdfdiiif")
```

---

## ahd 

Real data on Alzheimer's Disease. 

- Short URL: https://bit.ly/amlr-ahd
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/ahd.csv

```{r, echo = FALSE, message=FALSE}
read_csv("https://bit.ly/amlr-ahd", col_types = list(male = "f", Genotype = "f", diagnosis = "f"))
```

---

## hcp_memory

Real neuroimaging data from the Human Connectome Project used to predict scores on a memory test. Note: artificially modified to increase predictive power and make activities more engaging.  

- Short URL: https://bit.ly/amlr-memory
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/hcp_memory.csv

```{r, echo = FALSE, message=FALSE}
read_csv("https://bit.ly/amlr-memory")
```

---

## water

Real data on 3276 different water bodies. Modified to turn `Potability` from a numeric variable (dummy code) into a character variable.

- Short URL: https://bit.ly/amlr-water
- Long URL: https://github.com/PittMethods/appliedml/blob/main/data/water.csv
- Original Source: https://www.kaggle.com/adityakadiwal/water-potability

```{r, echo=FALSE, message=FALSE}
read_csv("https://bit.ly/amlr-water")
```

---

## attrition

Simulated data on 1470 fictional employees who either quit their job (attrition = yes) or did not (attrition = no).

- Available from the {modeldata} package

```{r, echo=FALSE, message=FALSE}
data("attrition", package = "modeldata")
attrition
```

