---
title: '<span style="font-size:48pt;">RF Example and Reporting</span>'
subtitle: '🌲  ✍️   📝 '
author: 'Pittsburgh Summer Methodology Series'
date: 'Day 3D &emsp; &emsp; August 10, 2022'
output:
  xaringan::moon_reader:
    css: [../css/xaringan-themer.css, ../css/styles.css]
    nature:
      slideNumberFormat: "%current% / %total%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
      navigation:
        scroll: false
    self_contained: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  fig.showtext = TRUE,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE,
  collapse = TRUE
)
```

```{r xaringanExtra, echo=FALSE}
xaringanExtra::use_tachyons()
xaringanExtra::use_clipboard()
xaringanExtra::use_tile_view()
```

```{r packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(countdown)
library(patchwork)
library(xaringanthemer)
library(tidymodels)
tidymodels_prefer()
```

class: inverse, center, middle
# Random Forest Example

---

## Live Coding: Prepare packages and data

```{r, eval=FALSE}
# Load packages

library(tidyverse)
library(tidymodels)
tidymodels_prefer()

# Install model package (if needed)

install.packages("ranger")

# Load and view data

water <- read_csv("https://tinyurl.com/water-pm")
water
```

---

## Live Coding: Split data

```{r, eval=FALSE}
# Set random number generation seed for reproducibility

set.seed(2022)

# Create initial split for holdout validation

pot_split <- initial_split(water, prop = 0.8, strata = Potability)
pot_train <- training(pot_split)
pot_test <- testing(pot_split)

# Create 10-fold CV within training set for tuning

pot_folds <- vfold_cv(pot_train, v = 10, repeats = 1, strata = Potability)
```

---

## Live Coding: Set up workflow

.scroll.h-0l[
```{r, eval=FALSE}
# Set up recipe based on the data and model

pot_recipe <-
  recipe(pot_train, formula = Potability ~ .) %>%
  step_mutate(Potability = factor(Potability)) %>%
  step_nzv(all_predictors()) %>%
  step_corr(all_numeric_predictors()) %>%
  step_lincomb(all_numeric_predictors()) %>%
  step_normalize(all_predictors())

# Set up model with tuning parameters and variable importance configuration

rf_model <-
  rand_forest(mtry = tune(), trees = tune(), min_n = tune()) %>%
  set_mode("classification") %>%
  set_engine("ranger", importance = "impurity")

# Combine recipe and model specification into a workflow

pot_wflow <-
  workflow() %>%
  add_recipe(pot_recipe) %>%
  add_model(rf_model)
```
]

---

## Live Coding: Set up and run tuning

```{r, eval=FALSE}
# Pick reasonable boundaries for the tuning parameters

pot_param <-
  rf_model %>%
  extract_parameter_set_dials() %>%
  finalize(pot_folds)

# Create list of values within boundaries and grid search (may take a bit)

pot_tune <-
  pot_wflow %>%
  tune_grid(
    resamples = pot_folds,
    grid = 10,
    param_info = pot_param
  )

# If desired, plot the tuning results

autoplot(pot_tune)
```

---

## Live Coding: Finalize the workflow

```{r, eval=FALSE}
# Select the best parameters values

pot_param_final <- select_best(pot_tune, metric = "roc_auc")
pot_param_final

# Finalize the workflow with the best parameter values

pot_wflow_final <-
  pot_wflow %>%
  finalize_workflow(pot_param_final)
pot_wflow_final

# Fit the finalized workflow to the training set and evaluate in testing set

pot_final <-
  pot_wflow_final %>%
  last_fit(pot_split)
```

---

## Live Coding: Explore performance

.scroll.h-0l[
```{r, eval=FALSE}
# View the metrics (from the holdout test set)

collect_metrics(pot_final)

# Collect the predictions

pot_pred <- collect_predictions(pot_final)
pot_pred

# Calculate and plot confusion matrix

pot_cm <- conf_mat(pot_pred, truth = Potability, estimate = .pred_class)
pot_cm

autoplot(pot_cm, type = "mosaic")
autoplot(pot_cm, type = "heatmap")

summary(pot_cm)

# Plot the predicted class probabilities

ggplot(pot_pred, aes(x = .pred_safe, y = Potability)) + 
  geom_boxplot()

# Plot the ROC curve

pot_rc <- roc_curve(pot_pred, truth = Potability, estimate = .pred_safe)

autoplot(pot_rc)
```
]

---

## Live Coding: Interpretation

```{r, eval=FALSE}
# Plot the variable importance measure (requires setting importance earlier)

library(vip)
pot_final %>%
  extract_fit_parsnip() %>%
  vip()
```







---

class: inverse, center, middle
# Reporting

---

class: inverse, center, middle

# End of Day 3
