---
title: '<span style="font-size:48pt;">Practical Issues</span>'
subtitle: '.big[🎨 🔬 📝]'
author: 'Applied Machine Learning in R </br> Pittsburgh Summer Methodology Series'
date: 'Lecture 4-B &emsp; &emsp; July 22, 2021'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      beforeInit: ["macros.js", "cols_macro.js"]
      slideNumberFormat: "%current% / %total%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE,
  par(bg = "#E9ECEF")
)
knitr::opts_knit$set(global.par = TRUE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#23395b",
  footnote_font_size = "20px",
  footnote_color = "gray",
  text_slide_number_font_size = "18px"
)
```

```{r packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(countdown)
library(patchwork)
library(summarytools)
library(mice)
library(VIM)
library(caret)
library(gridExtra)
library(psych)
library(psychTools)
library(ggcorrplot)
library(GGally)
library(corrplot)
library(AppliedPredictiveModeling)
```

class: inverse, center, middle
# Overview

---
## Lecture Topics

.pull-left[
**Planning Studies**
- Identifying interesting questions
- Sample size

**Data Quality** 
- Measurement error
- Imbalanced datasets
- Bias, representativeness, and generalizability

**Interpreting Results**
- When to trust your model 
- Data-generating mechanism
- External validation
- Robustness checks
]

.pull-right[
**Writing**
- Information to include in Methods
- Information to include in Results

**Publication Tables and Figures**
- ggplot? 

**Peer-Reviewing ML Papers**
- Sample 
- Measurement
- Is ML necessary? 
- Alignment of data, research question, methods, and algorithms 
]

---
class: inverse, center, middle
# Planning Studies 

---
## Planning Studies


---
class: inverse, center, middle
# Data Quality 

---
## Data Quality 


---
class: inverse, center, middle 
# Interpreting Results


---
## Interpreting Results


---
class: inverse, center, middle
# Writing  

---
## Writing Methods


---
## Writing Results


---
class: inverse, center, middle
# Publication Tables and Figures

---
## Publication Tables and Figures


---
class: inverse, center, middle
# Peer-Reviewing ML Papers


---
## Peer-Reviewing ML Papers


