---
title: '<span style="font-size:48pt;">Support Vector Machines</span>'
subtitle: '🚧  🛣️  🦿 ' 
author: 'Applied Machine Learning in R <br />Pittsburgh Summer Methodology Series'
date: 'Lecture 4-A &emsp; &emsp; July 22, 2021'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      beforeInit: "macros.js"
      slideNumberFormat: "%current% / %total%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
    self_contained: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  fig.showtext = TRUE,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE,
  comment = "#>",
  collapse = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#23395b",
  footnote_font_size = "20px",
  footnote_color = "gray",
  text_slide_number_font_size = "18px"
)
```

```{r xaringan-tachyons, echo=FALSE}
xaringanExtra::use_tachyons()
```

```{r packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(countdown)
library(patchwork)
library(caret)
library(recipes)
```

class: inverse, center, middle
# Overview

<style type="text/css">
.onecol {
    font-size: 26px;
}
.twocol {
  font-size: 24px;
}
.remark-code {
  font-size: 24px;
  border: 1px solid grey;
}
a {
  background-color: lightblue;
}
.remark-inline-code {
  background-color: white;
}
</style>
---
## Maximal margin classifier

---
## Support vector classifier

---
## Support vector machine

---
## Support vector regression

---
# Kernels

---
## Overview of kernerls

---
## Polynomial kernel

---
## Radial basis function (RBF) kernel

---
## Hyperparameters for SVM

---
## Applied Example

---
## Live Coding Activity

---
## Hands-on Activity

---
## Break and timer
