---
title: "Day 2-A Activity <br /> (Feature Engineering)"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    df_print: paged
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  df_print = "paged"
)
```

# Load packages

```{r, message=FALSE}
library(tidyverse)
library(skimr)
library(caret)
library(recipes)
```

---

# Read in data from CSV file 

```{r, message=FALSE}
airsat <- read_csv("https://bit.ly/amlr-airsat")
airsat
```

---

# Create training and testing sets

```{r}
set.seed(2021)
index <- createDataPartition(airsat$satisfaction, p = 0.8, list = FALSE)
airsat_train <- airsat[index, ]
airsat_test <- airsat[-index, ]

dim(airsat_train)
dim(airsat_test)
```

---

# Explore the training set

```{r}
skim(airsat_train)
```

---

# Create and prep recipe

```{r}
airsat_recipe <- 
  airsat %>% 
  recipe(satisfaction ~ .) %>% 
  step_nzv(all_predictors()) %>% 
  step_lincomb(all_numeric_predictors()) %>% 
  step_normalize(all_numeric_predictors()) %>% 
  step_pca(all_numeric_predictors(), threshold = 0.9) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  prep(training = airsat_train, log_changes = TRUE)
```

---

# Bake new training set

```{r}
airsat_baked_train <- bake(airsat_recipe, new_data = airsat_train)
airsat_baked_train
```

---

# Bake new testing set

```{r}
airsat_baked_test <- bake(airsat_recipe, new_data = airsat_test)
airsat_baked_test
```

---

# Hands-on Activity

Modify the code above to accomplish the following goals:

 1. Use 75% of the data for training and 25% of the data for testing.

 1. Apply the Yeo-Johnson transformation to the `flight_distance` variable (before normalizing it).

 1. Instead of using PCA to address multicollinearity, drop highly correlated predictors.

 1. Use one-hot encoding for the nominal predictors instead of dummy codes.

 1. Add an interaction term that conditions `seat_comfort` on `flight_distance`

**BONUS:** Read the "Recommended preprocessing" appendix (https://www.tmwr.org/pre-proc-table.html) 

**FURTHER READING:** https://www.tmwr.org/recipes.html (Chapter) https://bookdown.org/max/FES/ (Book)

---

## Answer key

<details><summary>Click here to view the answer key to the hands-on activity</summary>

```{r}
set.seed(2021)
index <- createDataPartition(airsat$satisfaction, p = 0.75, list = FALSE) #1
airsat_train <- airsat[index, ]
airsat_test <- airsat[-index, ]
```

```{r}
airsat_recipe <- 
  airsat %>% 
  recipe(satisfaction ~ .) %>% 
  step_nzv(all_predictors()) %>% 
  step_lincomb(all_numeric_predictors()) %>% 
  step_YeoJohnson(flight_distance) %>%                      #2
  step_normalize(all_numeric_predictors()) %>% 
  step_corr(all_numeric_predictors()) %>%                   #3 
  step_dummy(all_nominal_predictors(), one_hot = TRUE) %>%  #4 
  step_interact(~ seat_comfort:flight_distance) %>%         #5
  prep(training = airsat_train, log_changes = TRUE)
```

</details>
