---
title: '<span style="font-size:48pt;">Basic Predictive Modeling</span>'
subtitle: '🧮 🔮 👨‍💻' 
author: 'Applied Machine Learning in R <br />Pittsburgh Summer Methodology Series'
date: 'Lecture 2-B &emsp; &emsp; July 20, 2021'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      beforeInit: "macros.js"
      slideNumberFormat: "%current% / %total%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  fig.showtext = TRUE,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_light(
  base_color = "#23395b",
  footnote_font_size = "20px",
  footnote_color = "gray",
  text_slide_number_font_size = "18px"
)
```

```{r packages, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(countdown)
library(patchwork)
library(caret)
library(recipes)
```

class: inverse, center, middle
# Overview

<style type="text/css">
.onecol {
    font-size: 26px;
}
.twocol {
  font-size: 24px;
}
</style>
---
class: onecol
## Lecture Topics
In this lecture, we will discuss 


---

Training and Tuning


Performance Evaluation
-

Variable importance


---
class: inverse, center, middle
# Training and splitting
---



---
class: inverse, center, middle
# Hyperparameter tuning
---
class: inverse, center, middle
# Performance evaluation
---
class: inverse, center, middle
# Variable Importance

